Dealing with Consumer Uncertainty: Online Consumer Reviews and What makes them helpful?

### 2022-11-07 Paper.Rmd
many fixes :(

---
documentclass: report
title: 'Dealing with Consumer Uncertainty: Online Consumer Reviews and What makes them helpful?'
bibliography: references.bib
csl: apa.csl
link-citations: yes
output:
  bookdown::pdf_document2:
    fig.caption: yes
    includes:
      in_header: [non-float-fig.tex, custom.tex]
    toc_depth: 4
geometry: left=40mm, right=20mm, top=30mm, bottom=30mm
fontsize: 12pt
papersize: A4
linestretch: 1.5
fontfamily: times
---

@@

Online shopping has becoming more and more prevalent in average consumer's daily life. Electronic shopping brought many advantages both for sellers and buyers. One of the positive perks is Online Consumer Reviews (OCR). Online Consumer Reviews come in the form of electronic word-of-mouth [@dellarocas2003] and can be defined as "*peer-generated product evaluations posted on company or third party websites*" [@mudambi2010]. OCR have an effect on product sales as well as consumer decision making [@duan2008; @lee2011; @dellarocas2007].

OCR help customers by decreasing their search costs. Customers try to get information on the product they are going to buy so that they can decrease the uncertainty about that particular product. This information acquiring process is costly. Furthermore, this information gathering process might be more or less costly depending on the product type, namely Search and Experience goods [@nelson1970] . Search goods, by nature, have traits that can be observed objectively before buying (e.g., GHz of a CPU of a Laptop). On the other hand, quality of Experience goods can not be evaluated before buying and evaluations could be very subjective (e.g., One needs to watch a movie in order to understand if it is a good movie or not.).

However, the increased rate of electronic commerce activity also created an abundance of product reviews on the internet. The vast amount of product reviews might increase the information overload on the customer [@sun2019; @fan2021]. Unfortunately, this information overload can make the search process even costlier because it is really time consuming to evaluate thousand of consumer reviews. Additionally, not all reviews have the same informativeness level [@pan2011]. Therefore, answering the question of "what makes a helpful review?" is very important because reading more and more reviews without getting useful information can only increase the search cost. To make the understanding of a "helpful review" more clear it can be defined as "*a peer-generated product evaluation that facilitates the consumer's purchase decision process*" [@mudambi2010].

Having a good understanding of how consumers handle information and how their behaviors change with respect to information they are exposed to might even unearth newer tools to increase the social welfare because it would lead to decreased level of information asymmetry. The decreased information asymmetry might produce two benefits. First, consumers can learn to evaluate the quality of the products before purchase much more efficiently. This leaves less room for any dis-utility that can arise from purchasing a sub-par quality product. The second, the increased knowledge of the consumers towards certain products may even facilitate competition among firms. This competition may lead increased quality of the products as well as decreased prices in a given market. This is why understanding the determinants of a helpful review is important because it might reveal new insights on consumer behavior and these new insights could be used in favor of both consumers are producers.

Since OCR do not share the same level of informativeness and usefulness [@pan2011], most of the electronic retail websites implemented a very well known "helpfulness" feature to their customer review systems in order to help their customers navigate through thousand of reviews. On some websites, this feature is implemented in a way that helpful / unhelpful buttons are present and on some other websites, there is only a helpful button. In this study, the data was acquired from websites which support helpful / unhelpful system since it was found that helpfulness as a percentage (helpful votes over total votes) creates more powerful results than considering only helpful votes as a count variable [@ahmad2015].

The research on perceived helpfulness is still growing. In Social and Behavioral Sciences, it is common to have conflicting papers and results in the literature [@hunter1991]. In this case, the study on perceived helpfulness is not an exception. Because of this reason, more empirical studies on perceived helpfulness are required. This paper aims to add one more perspective to the growing literature of customer reviews and review helpfulness by employing a unique data set as well as questioning the subject with the intuition of economics. Furthermore, this paper tries to show that unstructured data (e.g., text, image, etc.) is viable for economic research. With unstructured data, economics may be able to explore new venues.

@@

The plan of this study is outlined as the following. In the next section, the literature on the subject of Review Helpfulness is going to be presented. In the second section, the related theories which would help with the understanding of the consumer behavior and review helpfulness will be laid out. Next, the data and the research methodology are going to be introduced. The results will be shared in the fourth section. In the last section the results are going to be discussed and then the study is going to be concluded.

@@

The research on Online Consumer Reviews could be considered to have two main branches. One branch is heavily marketing oriented and tries to understand the effects of OCR on sales [@ghose2011; @ögüt2012]. The other branch is concerned with rather behavioral or psychological outcomes and focuses on perceived helpfulness only. The boundary between these two branches are somewhat blurry and the same independent variables were used to explain both helpfulness and sales. Furthermore, @topaloglu2019 found that helpfulness have a moderating effect on sales. Many other papers used perceived helpfulness to explain sales as well [@ghose2011; @lee2018]. Thus, these two branches are inter-connected with each other. This paper tries to understand behavioral aspect of the review helpfulness rather than its effects on sales.

Many independent variables and their relationship between perceived helpfulness were examined by researchers. @hong2017 groups these independent variables into two chunks. The first group is reviewer-related predictors. Variables within this group try to understand the relationship between reviewer status (reviewer expertise, number of past reviews etc.) and perceived helpfulness. [@sun2019; @ghose2011]. The second group tries to understand the review-related relationships such as review length, review extremity, discrete review emotion etc [@mudambi2010; @ghose2011; @ren2019; @eslami2018]. @ghose2011 showed that review-related predictors have stronger prediction power than reviewer-related ones. The choice of the variables is heavily dependent on the data itself. In my paper, only review-related predictors were examined because the data which is available for this study only allows for review-related variables.

Review Length (Depth) could be considered as one of the most important independent variable explaining perceived helpfulness. It might be implemented in different ways but for the majority of the studies it is the word count in a given review. For most of the researchers, longer reviews contain more information about the product. As the potential buyer gets more information from a single review, he decreases his search costs with minimum effort [@mudambi2010; @hong2017; @liu2014; @wang2018; @korfiatis2012; @lee2016; @lee2014; @choi2020]. However, other group of researchers found out that there was a threshold on review length, meaning longer reviews might have useless information in them[@schindler2012; @huang2015]. Another group indicates the word count of the review shows diminishing returns to helpfulness, supporting the threshold idea [@baek2012; @kuan2015; @eslami2018]. Although very rare, a negative relationship between Review Length and Perceived Helpfulness was also reported [@chatterjee2020; @yang2019].

Review Rating (Review Extremity) is another important identifier of perceived helpfulness. This is usually the star rating on a 1-5 or 1-10 scale and resembles the evaluation of the reviewer for a given product. Its effect on perceived helpfulness is rather divergent in the literature. A number of researchers found out that there is a non-linear relationship between review extremity and perceived helpfulness. Specifically, a U-Shaped function which means that reviews that are on the extremes (either very negative or very positive) are more helpful than neutral reviews[@mudambi2010; @cao2011; @liu2014; @wang2018; @ghose2006]. In other studies, linear positive relationship was found. In other words, positively rated reviews are more useful [@pan2011; @schindler2012; @korfiatis2012; @lee2016; @choi2020]. There are also results that support negative linear relationship[@chua2014] and no effect at all[@hong2017].

Review Age is an indicator of how long the time passed after a review published on a web site. On some papers the opposite of Review Age, Review Freshness (Recency) could be employed as well. Different effects on perceived helpfulness were also observed by researchers. Some researchers discussed that a timely review is very important at conveying the true information about a certain product [@liu2014; @li2019; @zhou2019]. Others discussed that some web sites promotes newer reviews and because of that more recent reviews are more helpful [@lee2014; @choi2020; @wu2021].

The counts of negative or positive words out of total word count (the proportion of negative and positive words) were employed by a handful of studies [@baek2012; @kuan2015]. These metrics can measure the sentiment of the review itself. They are usually different than Review Extremity. In most cases Review Extremity and Review Sentiment (Valence) do not match [@zhang2011]. @baek2012 and @kuan2015 discovered a positive relationship between the proportion of negative words and perceived helpfulness. The relationship between the proportion of positive words and perceived helpfulness was found to be negative [@kuan2015].

@@

Many different types of research methodologies were employed in the literature of the Review Helpfulness. To my knowledge the most common empirical analysis method that was used in the literature was Tobit Regression because of the limited and censored nature of the Reviews Helpfulness [@mudambi2010; @korfiatis2012; @huang2015; @choi2020; @ren2019; @chua2014]. Some of these studies also shared Ordinary Least Squares results alongside with the Tobit Regression results for the purpose of benchmarking [@mudambi2010; @kuan2015; @choi2020]. Ordinary Least Squares were also used as a main empirical methodology in some papers [@liu2014; @ghose2006; @lee2016]. Furthermore, Ordinary Least Squares was employed in a hierarchical manner as well [@baek2012]. Some other studies employed Multi-Level Regression models in order to deal with potential biases [@ghose2011; @wu2021; @kuan2015]. In a rare case Ordinal Logistic Regression was one of the methods that was used to explain the determinants of the review helpfulness [@cao2011]. Predictive methods such as Neural Networks were used to understand the components of a helpful review as well [@lee2014]. One paper attacked the problem of the review helpfulness with three methods, namely, Partial Least Squares - Structural Equation Modelling, ANOVA Analysis and Artificial Neural Networks [@eslami2018].

@@

Since there are many conflicting results in the literature of review helpfulness additional empirical contributions are important. As well as adding new empirical findings to the literature, this study tries to extend the potential predictors with the interaction term between review length and review rating to test for the existence of the Negativity Bias. Furthermore, some under-utilized predictors from the literature such as the proportion of negative words and the proportion of positive words are also used to test Negativity Bias and Loss Aversion.

@@

Relaxing the assumption that consumers have perfect information about the products on the market raises the problem of information acquisition for consumers. Getting the complete information about products is important because economic agents want to maximize their utility by finding the best good on the market. The complete information is the information about both the prices and quality levels of the goods which are available on the market. Acquiring the information of the quality of a product is harder than getting the information about its price. To explain, a consumer can check the price list of laptops and have an idea about maximum and minimum prices but getting the information about a laptop's technical components or a brand's support for its products is a lot more effort taking and costly.

@@

Consumers can choose from two different strategies when they want to acquire more information about products. Consumers "search" in order to get information about a product assuming that they perfectly know where they can obtain the knowledge. @nelson1970 defines the search process as *"The most obvious procedure available to the consumer in obtaining information about price or quality is search"*. The search process is complete when consumers inspect the option and that inspection happens before the purchase. Contrary to the search process, consumer may choose to "experience" the product. The consumer follows this strategy when the search for information is either inappropriate, inefficient or costly compared to just experiencing the good. This is true for cheap goods especially. For instance, in order to find out the best chocolate in the market it is best for the consumer to buy every other brand available and taste them. Thus, consumers will always prefer sampling or trial use given the fact that the cost of the sampling or trial is lower than the cost of the search. This is why search is costlier for Experience goods. Therefore, the kind of strategy that a consumer chooses depends on the product's Search or Experience attributes and how dominant these attributes are.

@nelson1970 also states that consumers can use other consumers' experiences in their search processes. The information about a certain good is usually acquired in the form of "word-of"mouth" in this kind of search process. This type of search is viable when getting the information about a certain product is much more costly by "experience" compared to by "search". @klein1998 discusses that experience sharing becomes easier with the increasing usage of the internet. Because experiences about products can be shared much more easily on the internet, consumers' pre-purchase behavior might change accordingly. The reason of this is that internet decreases search costs considerably. Furthermore, the potential consumer may engage in "vicarious learning" where he/she gets indirect usage experience through the experiences of other consumers in the search process [@murray1991].

Ultimately, to put the theory in a more generalized way, a product's Total Cost consists of both Search Cost and Product Cost. This Search Cost is higher for Experience goods, in other words, information search is costlier for Experience goods than Search goods. In the consumer's perspective, to make the search viable, the marginal benefit of an additional search must be grater than the marginal cost of the additional search. The increasing use of internet decreases the search costs. Because of the decreased search costs, the consumer may undertake more search processes and this might change how people perceive good types, in other words, internet and virtual experience sharing can convert some Experience goods to Search goods [@klein1998].

@@

Negativity Bias Theorem argues that humans or even animals have tendencies to be more aware of negative entities than positive ones. In @rozin2001 's words Negativity Bias can be best exemplified such as following:

> *"Brief contact with a cockroach will usually render a delicious meal inedible. The inverse phenomenon rendering a pile of cockroaches on a platter edible by contact with one's favorite food is unheard of. More modestly, consider a dish of a food that you are inclined to dislike: lima beans, fish, or whatever. What could you touch to that food to make it desirable to eat that is, what is the anti-cockroach? Nothing! And the cockroach is far from unique: there is a wide variety of animals (e.g., worms, caterpillars, slugs, spiders) that share the cockroach potency, along with a variety of microbially or toxin-contaminated objects. One of the best generic descriptions of this relative power of negative contamination is embedded in an age old Russian adage: 'A spoonful of tar can spoil a barrel of honey, but a spoonful of honey does nothing for a barrel of tar.'"*

@@

**Negative Potency** defends that the exact opposite of something positive carries heavier weight. In other words, negative events are much more impactful than their exact positive counterparts. For instance, losing 100\$ could have a harder psychological effect on a person than that particular person is gaining 100\$. Loss Aversion [@kahneman1979; @tversky1991] is the phenomenon where the Negative Potency can be observed the best. People risk more (less) when they are losing (gaining) something. This behavior can be explained by the Endowment Effect [@kahneman1990]. Endowment effect occurs because people usually over-estimate the value of their belongings, for example a pencil. However, they would have under-valued the same pencil if the item did not belong to them. The reason to this behavior is that the dis-utility of losing that pencil is greater than the utility of acquiring the same pencil.

**Greater Steepness of Negative Gradients** simply implies that the marginal psychological effect of a negative entity is much more higher than a positive one. In other words, negative things cause more negativity as the time goes on. At the very least, the effects of negativeness are a lot more long-standing than effects of positiveness.

@@

> *''The image that comes to mind from the physical domain is that of a single cancerous growth or germ that radiates itself through and ultimately consumes a perfectly healthy body. The image that comes to mind from the moral domain is that of a single vice corrupting and perverting and bringing the moral downfall of an otherwise perfectly good person.''*

**Greater Negative Differentiation** discusses the possibility that human cognition is fine-tuned to negative entities because being aware of negative stimuli might correlate with survival in the process of the evolution. As a consequence, the vocabulary that reflects and describes the state of negativeness is much more wider than its counterpart positiveness. [@peeters1971].

@@

The data was scraped with a custom web scraper which was constructed with Selenium Framework for Python 3 [@python3]. The scraper got the customer reviews from IMDB and BestBuy. The former is the biggest platform for Movies and TV Shows, the latter is one of the biggest electronic retailers in United States of America. Both websites host a great amount of online reviews. After the data was collected it was further transformed and analyzed with the help of R Programming Language [@RCore; @Tidyverse; @Tidytext].

Online Consumer Reviews from IMDB were collected as an experience goods data since movies can be considered as pure-experience goods and their search attributes are very limited. It is nearly impossible to evaluate a movie before watching it. Furthermore, evaluations for a movie can both have subjective (e.g., personal attachment to the protagonist, atmosphere of the movie etc.) and objective aspects (e.g., writing, production quality etc.). There are also numerous studies that analyzed movies as experience goods[@nelson1970; @liu2014; @ghose2006; @baek2012; @lee2016].

BestBuy is a very big electronics retailer and this is why the website offers a sizable number of OCR for search goods. Contrary to experience goods, search goods usually have more objectively quantifiable traits. For example, resolution or screen size of a TV, RAM or GPU of a Laptop, storage or battery lifetime of a Smart Phone. Additionally, in the literature, TVs, Laptops and Phones were used as search goods [@mudambi2010; @sun2019; @baek2012; @ren2019]. Therefore, search goods data was collected from BestBuy.

From IMDB reviews of 22 movies were collected as experience goods. In total, 22 movies account for 31,863 reviews. From BestBuy, total of 383,357 reviews were collected as search goods. Specifically, 201,113 reviews for TVs, 130,270 for Laptops and 51,974 for Phones. However, in the data there are many reviews that were never got voted and probably they were never read because they got lost in the crowd of other reviews. To make sure that the analysis sample contains only the reviews which were actually read by potential buyers, a threshold of 5 was set for Total Votes. In other words, reviews that had less than 5 Total Votes were removed from the analysis data [@choi2020; @baek2012]. At the very end, the analysis data was consisted of 15,024 observations for experience goods and 8,042 observations for search goods. It should be noted that `r round(15024 / 31863, digits = 2) * 100` percent of the reviews had total votes of 5 or greater for experience goods, whereas only `r round(8042 / 383357, digits = 2)* 100` percent of the search goods were voted at least 5 times. Most of the search product reviews are not getting voted and even possible that they are not getting read at all. This might support the idea that people form more homogeneous opinions about search goods [@hong2014]. Once a helpful review is written most of the other reviews become redundant since other reviews most likely convey similar information. On the contrary, more attendance to reviews can be observed for experience goods (movies) since nearly half of them were voted at least five times. This might be because people form heterogeneous opinions about experience goods [@hong2014].

@@

The goal of this section is not "confirmatory" whatsoever. Both dependent and independent variables are going to be introduced with an emphasis on their quantitative features such as their distributions and potential transformations which can be applied to the variables in question. In addition to this, bivariate relations between independent and dependent variables will be plotted in order to get an initial idea about the potential effects along with similar findings from the literature. Thus, in a sense, EDA in this section does not carry out the purpose of discovery, rather it merely puts the data on the spot for the sake of transparency.

@@

Black vertical lines show the mean ratings for both product types. It is nice to see that averages are close to the neutral rating of 3. With this it can be made sure that reviews do not appeal to positive or negative side only.

@@

>>>>

---
documentclass: report
title: 'Dealing with Consumer Uncertainty: Online Consumer Reviews and What makes them helpful?'
bibliography: references.bib
csl: apa.csl
toc: no
link-citations: yes
output:
  bookdown::pdf_document2:
    fig.caption: yes
    includes:
      in_header: [non-float-fig.tex, custom.tex]
    toc_depth: 4
geometry: left=40mm, right=20mm, top=30mm, bottom=30mm
fontsize: 12pt
papersize: A4
linestretch: 1.5
fontfamily: times
---

@@

Online shopping has becoming more and more prevalent in average consumer's daily life. Electronic shopping brought many advantages both for sellers and buyers. One of the positive perks is Online Consumer Reviews (OCR). Online Consumer Reviews come in the form of electronic word-of-mouth [@dellarocas2003] and can be defined as "*peer-generated product evaluations posted on company or third party websites*" [@mudambi2010]. OCR were found to have an effect on product sales as well as consumer decision making process. [@duan2008; @lee2011; @dellarocas2007].

OCR help customers by decreasing their search costs because it is easier for the consumer to skim through many reviews on the internet than finding and asking people who actually bought the product in question. Consumers try to get information on the product they are going to buy so that they can decrease the uncertainty about that particular product. This information acquiring process (search) is costly. Furthermore, this information gathering process might be more or less costly depending on the product type, namely Search and Experience goods [@nelson1970]. Search goods, by nature, have traits that can be observed objectively before buying (e.g., GHz of a CPU of a Laptop). On the other hand, the quality of Experience goods can not be evaluated before buying and evaluations could be very subjective (e.g., One needs to watch a movie in order to understand if it is a good movie or not. Also a movie being good or bad might depend on who watched it because people have different tastes.).

However, the increased rate of electronic commerce activity also created an abundance of product reviews on the internet. The vast amount of product reviews might increase the information overload on the customer [@sun2019; @fan2021]. Unfortunately, this information overload can make the search process even costlier because it is really time consuming to evaluate thousands of consumer reviews. Besides, not all reviews have the same informativeness level [@pan2011]. Therefore, answering the question of "what makes a helpful review?" is very important because reading more and more reviews without getting useful information can only increase the search costs. On top of this, the overload of useless information might impair the decision making ability of the potential buyer. To make the understanding of a "helpful review" more clear it can be strictly defined as "*a peer-generated product evaluation that facilitates the consumer's purchase decision process*" [@mudambi2010].

Having a good understanding of how consumers handle information and how their behaviors change with respect to information they are exposed to might even unearth newer tools to increase the social welfare because it would lead to a decreased level of information asymmetry between consumers and producers. The decreased information asymmetry might produce two benefits. First, consumers can learn to evaluate the quality of the products before purchase much more efficiently. This leaves less room for any dis-utility that can arise from purchasing a sub-par quality product. The second, the increased knowledge of the consumers towards certain products may even facilitate competition among firms. This competition may lead increased quality of the products as well as decreased prices in a given market. This is why understanding the determinants of a helpful review is important because it might reveal new insights on consumer behavior and these new insights could be used in favor of both consumers are producers.

Since OCR do not share the same level of informativeness and usefulness [@pan2011], most of the electronic retail websites implemented a very well known "helpfulness" feature to their customer review systems in order to help their customers navigate through thousand of reviews. On some websites, this feature is implemented in a way that *helpful* / *unhelpful* buttons are present and on some other websites, there is only a *helpful* button. In this study, the data was acquired from websites which support *helpful* / *unhelpful* system since it was found that helpfulness as a percentage (helpful votes over total votes) creates more powerful results than considering only helpful votes as a count variable [@ahmad2015].

The research on perceived helpfulness is still growing. In Social and Behavioral Sciences, it is common to have conflicting studies and results in the literature [@hunter1991]. In this case, the study on perceived helpfulness is not an exception. Because of this reason, more empirical studies on perceived helpfulness are required. This study aims to add one more perspective to the growing literature of customer reviews and review helpfulness by employing a unique data set as well as questioning the subject with the intuition of economics. Furthermore, this study tries to show that unstructured data (e.g., text, image, etc.) is a viable option for economic research. With unstructured data, economics may be able to explore new venues.

@@

The plan of this study is outlined as the following. In the next section, the literature on the subject of Review Helpfulness is going to be presented. In the second section, the related theories which would help with the understanding of the consumer behavior and review helpfulness will be laid out. Next, the data and the research methodology are going to be introduced. The results will be shared in the fourth section. In the fifth section the results are going to be discussed and then the study is going to be concluded in the last section.

@@

The research on Online Consumer Reviews can be considered to have two main branches. One branch is heavily marketing oriented and tries to understand the effects of OCR on sales [@ghose2011; @ögüt2012]. The other branch is concerned with rather behavioral or psychological outcomes and focuses on perceived helpfulness only. The boundary between these two branches are somewhat blurry and the same independent variables were used to explain both helpfulness and sales. Furthermore, @topaloglu2019 found that review helpfulness have a moderating effect on sales. Many other papers used the perceived helpfulness to explain sales as well [@ghose2011; @lee2018]. Thus, these two branches are inter-connected with each other. Contrary to sales oriented papers which were mentioned above, this study tries to understand behavioral aspects of the review helpfulness rather than its effects on sales.

Many different independent variables and their relationship between perceived helpfulness were examined by researchers. @hong2017 groups these independent variables into two chunks. The first group is reviewer-related predictors. Variables within this group try to understand the relationship between reviewer status (reviewer expertise, number of past reviews etc.) and perceived helpfulness. [@sun2019; @ghose2011]. The second group tries to understand the review-related relationships such as review length, review extremity, discrete review emotion etc. [@mudambi2010; @ghose2011; @ren2019; @eslami2018]. @ghose2011 showed that review-related predictors have stronger prediction power than reviewer-related ones. It should be noted that these two groups are not mutually exclusive and can be employed in the same study. The choice of the variables is heavily dependent on the data itself. In this particular study, only review-related predictors were examined because the data which is available for this study only allows for review-related variables. Reviewer-related variables could not be employed because IMDB and BestBuy use different conventions to identify their users(reviewers).

Review Length (Depth) could be considered as one of the most important independent variable explaining the perceived helpfulness. It might be implemented in different ways but for the majority of the studies it is the word count in a given review. For the most of the researchers, longer reviews contain more information about the product. As the potential buyer gets more information from a single review, he decreases his search costs with minimum effort [@mudambi2010; @hong2017; @liu2014; @wang2018; @korfiatis2012; @lee2016; @lee2014; @choi2020]. However, other group of researchers found out that there was a threshold on review length, meaning unnecessarily long reviews might have useless information in them[@schindler2012; @huang2015]. Another group indicates the word count of the review shows diminishing returns to helpfulness, supporting the threshold idea [@baek2012; @kuan2015; @eslami2018]. Although very rare, a negative relationship between Review Length and Perceived Helpfulness was also reported [@chatterjee2020; @yang2019].

Review Rating (Review Extremity) is another important identifier of perceived helpfulness. This is usually the star rating on a 1-5 or 1-10 scale and resembles the evaluation of the reviewer for a given product. Its effect on perceived helpfulness is rather divergent in the literature. A number of researchers found out that there is a non-linear relationship between review extremity and perceived helpfulness. Specifically, a U-Shaped function which means that reviews that are on the extremes (either very negative or very positive) are more helpful than neutral reviews [@mudambi2010; @cao2011; @liu2014; @wang2018; @ghose2006]. In other studies, linear positive relationship was found. In other words, positively rated reviews are more useful [@pan2011; @schindler2012; @korfiatis2012; @lee2016; @choi2020]. There are also results that support negative linear relationship [@chua2014] and no effect at all [@hong2017].

Review Age is an indicator of how long the time passed after a review published on a web site. In some papers the opposite of Review Age, Review Freshness (Recency) could be employed as well. Different effects on perceived helpfulness were also observed by researchers. Some researchers discussed that a timely review is very important at conveying the true information about a certain product, making older reviews more helpful [@liu2014; @li2019; @zhou2019]. Others discussed that some web sites promote newer reviews and recent information is much more valuable than old information. Because of these more recent reviews can be considered as more helpful [@lee2014; @choi2020; @wu2021].

The counts of negative or positive words out of total word count (the proportion of negative and positive words) were employed by a handful of studies [@baek2012; @kuan2015]. These metrics can measure the sentiment of the review itself. They are usually different than Review Extremity (Review Rating). In most cases Review Extremity and Review Sentiment (Valence) do not match [@zhang2011]. @baek2012 and @kuan2015 discovered a positive relationship between the proportion of negative words and perceived helpfulness. The relationship between the proportion of positive words and perceived helpfulness was found to be negative [@kuan2015].

@@

Many different types of research methodologies were employed in the literature of the Review Helpfulness. To my knowledge the most common empirical analysis method that was used in the literature was Tobit Regression because of the limited and censored nature of the Reviews Helpfulness [@mudambi2010; @korfiatis2012; @huang2015; @choi2020; @ren2019; @chua2014]. Some of these studies also shared Ordinary Least Squares results alongside with the Tobit Regression results for the purpose of benchmarking [@mudambi2010; @kuan2015; @choi2020]. Ordinary Least Squares were also used as a main empirical methodology in some papers [@liu2014; @ghose2006; @lee2016]. Furthermore, Ordinary Least Squares was employed in a hierarchical manner as well [@baek2012]. Some other studies employed Multi-Level Regression model in order to deal with the potential selection bias which is present in the data [@ghose2011; @wu2021; @kuan2015]. In a rare case Ordinal Logistic Regression was one of the methods that was used to explain the determinants of the review helpfulness [@cao2011]. Predictive methods such as Neural Networks were used to understand the components of a helpful review as well [@lee2014]. One paper attacked the problem of the review helpfulness with three methods, namely, Partial Least Squares - Structural Equation Modelling, ANOVA Analysis and Artificial Neural Networks [@eslami2018].

@@

Since there are many conflicting results in the literature of review helpfulness additional empirical contributions are important. As well as adding new empirical findings to the literature, this study tries to extend the potential predictors with the continuous interaction term between review length and review rating to test for the existence of the Negativity Bias. Furthermore, some under-utilized predictors from the literature such as the proportion of negative words and the proportion of positive words are also used to test Negativity Bias and Loss Aversion.

@@

Relaxing the assumption that consumers have perfect information about the products on the market raises the problem of information acquisition for consumers. Getting the complete information about products is important because economic agents want to maximize their utility by finding the best good with respect to their budget constraints on the market. The complete information is the information about both the prices and quality levels of the goods which are available on the market. Acquiring the information about the quality of a product is harder than getting the information about its price. To explain, a consumer can check the price list of laptops and have an idea about the maximum and the minimum prices but getting the information about a laptop's technical components or a brand's support for its products is a lot more effort taking and costly.

@@

Consumers can choose from two different strategies when they want to acquire more information about products. Consumers "search" in order to get information about a product assuming that they perfectly know where they can obtain the knowledge. @nelson1970 defines the search process as *"The most obvious procedure available to the consumer in obtaining information about price or quality is search"*. The search process is complete when consumers inspect the option and that inspection happens before the purchase. Contrary to the search process, consumer may choose to "experience" the product. The consumer follows this strategy when the search for information is either inappropriate, inefficient or costly compared to just experiencing the good. This is true for cheap goods especially. For instance, in order to find out the best chocolate in the market it is best for the consumer to buy every other brand available and taste them. Thus, consumers will always prefer sampling or trial use (experiencing the product) given the fact that the cost of the sampling or trial is lower than the cost of the search. This is why search is costlier for Experience goods. Therefore, the kind of strategy that a consumer chooses depends on the product's Search or Experience attributes and how dominant these attributes are.

@nelson1970 also states that consumers can use other consumers' experiences in their search processes. The information about a certain good is usually acquired in the form of "word-of-mouth" in this kind of search process. This type of search is viable when getting the information about a certain product is much more costly by "experience" compared to by "search". For instance buying a laptop and trying to figure out if it is the best laptop for a certain task may even require buying additional laptops because of the need for comparison. This is the way of "experiencing" the good and this would be very costly. On the other hand, the potential buyer may want to hear from people who do the similar tasks with their laptops and try to sort out the best possible laptop model for her own task. This way she gets information about the product without having to "experience" it and this approach is way cheaper.

@klein1998 discusses that experience sharing becomes easier with the increasing usage of the internet. Because experiences about products can be shared much more easily on the internet, consumers' pre-purchase behavior might change accordingly. The reason of this is that internet decreases search costs considerably. Furthermore, the potential consumer may engage in "vicarious learning" where he/she gets indirect usage experience through the experiences of other consumers in the search process [@murray1991].

Ultimately, to put the theory in a more generalized way, a product's Total Cost consists of both Search Cost and Product Cost. This Search Cost is higher for Experience goods, in other words, information search is costlier for Experience goods than Search goods. In the consumer's perspective, to make the search viable, the marginal benefit of an additional search must be grater than the marginal cost of the additional search. The increasing use of internet decreases the search costs. Because of the decreased search costs, the consumer may undertake more search processes and this might change how people perceive the good types, in other words, internet and virtual experience sharing can convert some Experience goods to Search goods [@klein1998].

@@

Negativity Bias Theorem argues that humans or even animals have tendencies to be more aware of negative entities than positive ones. In @rozin2001's words Negativity Bias can be best exemplified such as the following:

> *"Brief contact with a cockroach will usually render a delicious meal inedible. The inverse phenomenon rendering a pile of cockroaches on a platter edible by contact with one's favorite food is unheard of. More modestly, consider a dish of a food that you are inclined to dislike: lima beans, fish, or whatever. What could you touch to that food to make it desirable to eat that is, what is the anti-cockroach? Nothing! . . . One of the best generic descriptions of this relative power of negative contamination is embedded in an age old Russian adage: 'A spoonful of tar can spoil a barrel of honey, but a spoonful of honey does nothing for a barrel of tar.'"*

@@

**Negative Potency** defends that the exact opposite of something positive carries heavier weight. In other words, negative events are much more impactful than their exact positive counterparts. For instance, losing 300\$ could have a harder psychological effect on a person than that particular person is gaining 300\$. Loss Aversion [@kahneman1979; @tversky1991] is the phenomenon where the Negative Potency can be observed the best. People risk more (less) when they are losing (gaining) something. This behavior can be explained by the Endowment Effect [@kahneman1990]. Endowment effect occurs because people usually over-estimate the value of their belongings, for example a pencil. However, they would have under-valued the same pencil if the item did not belong to them. The reason to this behavior is that the dis-utility of losing that pencil is greater than the utility of acquiring the same pencil.

**Greater Steepness of Negative Gradients** simply implies that the marginal psychological effect of a negative entity is much more higher than a positive one. In other words, negative things cause more negativity as the time goes on. At the very least, the effects of the negativeness are a lot more long-standing than effects of the positiveness.

@@

> *"The image that comes to mind from the physical domain is that of a single cancerous growth or germ that radiates itself through and ultimately consumes a perfectly healthy body. The image that comes to mind from the moral domain is that of a single vice corrupting and perverting and bringing the moral downfall of an otherwise perfectly good person."*

**Greater Negative Differentiation** discusses the possibility that human cognition is fine-tuned to negative entities because being aware of negative stimuli might correlate with survival in the process of the evolution. As a consequence, the vocabulary that reflects and describes the state of the negativeness is much more wider than its counterpart the positiveness. [@peeters1971].

@@

The data was scraped with a custom web scraper which was constructed with Selenium Framework for Python 3 [@python3]. The scraper aqcuired the customer reviews from IMDB and BestBuy. The former is the biggest platform for Movies and TV Shows, the latter is one of the biggest electronic retailers in United States of America. Both websites host a great amount of online reviews. After the data was collected it was further transformed and analyzed with the help of R Programming Language [@RCore; @Tidyverse; @Tidytext].

Online Consumer Reviews from IMDB were collected as an experience goods data since movies can be considered as pure-experience goods and their search attributes are very limited. It is nearly impossible to evaluate a movie before watching it. Furthermore, evaluations for a movie can both have subjective (e.g., personal attachment to the protagonist, atmosphere of the movie etc.) and objective aspects (e.g., writing, production quality etc.). There are also numerous studies that analyzed movies as experience goods [@nelson1970; @liu2014; @ghose2006; @baek2012; @lee2016].

BestBuy is a very big electronics retailer and this is why the website offers a sizable number of OCR for search goods. Contrary to experience goods, search goods usually have more objectively quantifiable traits. For example, resolution or screen size of a TV, RAM or GPU of a Laptop, storage or battery lifetime of a Smart Phone. Additionally, in the literature, TVs, Laptops and Phones were used as search goods [@mudambi2010; @sun2019; @baek2012; @ren2019]. Therefore, search goods data was collected from BestBuy as it was a very suitable candidate for this research.

From IMDB reviews of 22 movies were collected as experience goods. In total, 22 movies account for 31,863 reviews. From BestBuy, total of 383,357 reviews were collected as search goods. Specifically, 201,113 reviews for TVs, 130,270 for Laptops and 51,974 for Phones. However, in the data there are many reviews that were never got voted and probably they were never read because they got lost in the crowd of other reviews. To make sure that the analysis sample contains only the reviews which were actually read by potential buyers, a threshold of 5 was set for Total Votes. In other words, reviews that had less than 5 Total Votes were removed from the analysis data [@choi2020; @baek2012]. At the very end, the analysis data was consisted of 15,024 observations for experience goods and 8,042 observations for search goods. It should be noted that `r round(15024 / 31863, digits = 2) * 100` percent of the reviews had total votes of 5 or greater for experience goods, whereas only `r round(8042 / 383357, digits = 2)* 100` percent of the search goods were voted at least 5 times. Most of the search product reviews are not getting voted and even possible that they are not getting read at all. This might support the idea that people form more homogeneous opinions about search goods [@hong2014]. Once a helpful review is written most of the other reviews become redundant since other reviews most likely convey similar information. On the contrary, more attendance to the reviews can be observed for experience goods (movies) since nearly half of them were voted at least five times. This might be because people form heterogeneous opinions about experience goods [@hong2014].

@@

The goal of this section is not "confirmatory" whatsoever. Both dependent and independent variables are going to be introduced with an emphasis on their quantitative features such as their distributions and potential transformations which can be applied to the variable in question. In addition to this, bivariate relations between independent and dependent variables will be plotted in order to get an initial idea about the potential effects along with similar findings from the literature. Thus, in a sense, EDA in this section does not carry out the purpose of discovery, rather it merely puts the data on the spot for the sake of transparency.

@@

Black vertical lines show the mean ratings for both product types. It is nice to see that averages are close to the neutral rating of 3. With this it can be made sure that the sample reviews do not address to positive or negative side only.

@@

#### Total Votes
